---
title: "HW4 MD CityCrime"
---

In this assignment, we will be working with a dataset of total crimes reported by Maryland cities for two consecutive years. You'll load the data, calculate the violent and property crime rates for each city for both years, and then calculate the percentage change in the crime rates between the two years. You'll also write a short story based on your findings.

### Task 1: Load the Tidyverse and Janitor

You will need to load the tidyverse library and the janitor library

```{r}
library(tidyverse)
library(janitor)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv. You will create a dataframe called `md_cities` that contains the data from this URL.

### Task 2: Load the Data

```{r}
md_cities <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv")
```
### Task 3: Calculating Rates

Calculate violent and property crime rates for each city for both years using `mutate()`. You'll create four new columns: `violent_rate_2019`, `property_rate_2019`, `violent_rate_2020`, and `property_rate_2020`. You'll need to figure out the proper unit rate, given that there are varying populations. There's no perfect choice; choose the one that works best for you. Save the results to a new dataframe called `md_cities_rates`.

```{r}
md_cities_rates <- md_cities |> mutate(violent_rate_2019 = violent2019/pop2019 * 1000000, property_rate_2019 = property2019/pop2019 * 1000000, violent_rate_2020 = violent2020/pop2020 * 1000000, property_rate_2020 = property2020/pop2020 * 1000000)
```

### Task 4: Describe the Rates

Examine the new dataframe, `md_cities_rates`, by clicking on it in the Environment tab in RStudio and sorting by the rate columns. Write a couple of sentences about the results, comparing and contrasting the cities with the highest rates (both violent and property) in 2019 and 2020. Do you see any patterns? What does this tell you about crime in Maryland cities? What's the story here?

After examining the rates, I have gathered that for violent rates in 2019 Baltimore City had the highest violent crime rates but in 2020, the violent crimes were the highest in Worcester, Ocean City which is a little interesting because Worcester much lower population. In 2019 the property rates were the highest in Worcester, Ocean City and in 2020 the property rates were the highest in Worcester, Ocean City. I believe the story here is that Worcester is having very high crime rates despite the fact that the population is not very high which poses the question, why is crime so high?


### Task 5: Calculate Percentage Change

Let's add two percentage change columns to your dataframe: one for violent crime rates and one for property crime rates. You'll calculate the percentage change between 2019 and 2020 for each city. Save the results to a new dataframe called `md_cities_change`.

```{r}
md_cities_change <- md_cities_rates |> mutate(violent_rate_2020 - violent_rate_2019/ violent_rate_2019 *100,
                                              property_rate_2019 - property_rate_2020/ property_rate_2020 *100)
```

### Task 7: Describe Percentage Changes

Look at the `md_cities_change` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change results. What's more interesting here, the patterns or the outliers? What's the story in the data?

From what I could gather from the new columns, it looks like Worcester is still the highest crime rate and Queenstown in Queen Anne's is extremely low on the range of percentage changes. I would say that the story in the data is still that Worcester has a very high crime rate.

### Task 8: Tell the Best Story

Write up a lede and nutgraf of a story with the _best_ story idea you've found in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

Worcester has the highest crime rate in Maryland for both property and violent crimes in 2020.

While Baltimore had the highest violent crime rate in 2019, Worcester, Ocean City had the highest crime rate the following year in 2020. Property-related crime rates, however were the highest in Worcester in both 2019 and 2020. What makes this data so intersting is the population difference between Baltimore and Worcester--with Baltimore having a population of 588,593 in 2020 while Worcester had a population of only 6,929 in 2020. 

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.

### Task 10: Using the 2020 violent crime rate, calculate the percentiles using `ntile()`. Save the results to a new dataframe called `md_cities_percentiles`. How many cities are above the 90th percentile? Calculate the Interquartile Range of the 2020 violent crime rate using `summarize()` and R's `IQR()` function. The syntax is similar to how `mean()` works. Do the results change your best story idea from the previous assignment? Why or why not?

```{r}
md_cities_percentiles <- md_cities_change |> mutate(percentile = ntile(violent_rate_2020, 100))
```

- There are 10 cities above the 90th percentile

```{r}
md_cities_percentiles |> summarize( IQR(violent_rate_2020))
```
- The results do not change the story idea from the previous assignment because Worcester, Ocean City is still at the highest percentile for crime which was similar to the other results in the data above. Due to the fact that the population is so low compared to the other cities like Baltimore, I think the fact that Worcester has such a low population and such a high crime rate is still extremely interesting.
